Comparing probabilistic predictive models applied to football

### 2015-05-17 template.tex
figura adicionada nas analises; revisei secao sobre scoring rules

@@

More precisely, a scoring rule is a function of two variables (or vectors, matrices etc) that assigns a real number (score) to each possible outcome $X$ and its respective prevision value, $P(X)$.
It is maximum for each $x$ when $x=P(X)$, and is nonincreasing both as $x$ departs from $P(X)$ (for each $P(X)$) and as $P(X)$ departs from $x$ (for each $x$).
Thus, the best score is achieved when knowledge of $X$ is asserted as the value of $X$ itself, whatever it might be. Most of the scoring rules also exhibit an arbitrary normalizing feature such as $S[x,P(X)]=0$ for each $x$.

Although many functions satisfy the criteria for a scoring rule specified above, not every scoring rule can be regarded as providing a fair assessment of the assertion $P(X)$, that is, not all scoring rules encourage honesty and accuracy in assessing numerical values for their previsions.
Those that do encourage a fair assessment are named \emph{proper scoring rules} \red{colocar referencia}.

Say a forecaster's uncertainty about $X$ is $P_0$. A proper scoring rule is a scoring rule $S$ such that the prevision value $P(X)$ that minimizes $E_{P_0}[S(X,P(X))]$ is $P_0$. 
In other words, if one receives a reward $S(x,P(x))$ when $x$ happens and if $S$ is a proper scoring rule, then the highest expected reward is obtained by reporting one's  true uncertainty about $X$.
The use of a proper scoring rule thus encourages the forecaster to be honest to maximize the expected reward.
Moreover, this method leads automatically to an overall comparison between the outcomes of different personal evaluations concerned with the same totality of events.
Therefore, the accumulated score may be considered a concrete measure of success of each forecaster or model.

@@

@@

The green surface in Figure \ref{fig:simplex} represents the 2-simplex, i.e. the set of points such that $p_1+p_2+p_3=1$ for non-negative values of $p_1$, $p_2$ and $p_3$.
The triangle representing the simplex has sides of length $\sqrt{2}$ and its height is $\sqrt{6}/2$.
Drawing a similar equilateral triangle with height 1 and sides $2\sqrt{3}/3$, it is possible to represent all the points of the simplex.
This new triangle is used to graphically display the predictions as internal points because the sum of the distances of every point inside it to each side, representing the probability of each event, is always one.

@@

\node[label={below:Home wins}] at (0,0) {};
\node[label={below:Draw}] at (1.1547,0) {};
\node[label={above:Home looses}] at (0.57735,1) {};

@@

\node[label={below:Home wins}] at (0,-0.05) {};
\node[label={below:Draw}] at (1.1547,-0.05) {};
\node[label={above:Home looses}] at (0.57735,1.05) {};

@@

\red{parei aqui}

\[p=(0.25,0.35,0.40)\]
\[d^2=(0-0.25)^2+(0-0.35)^2+(1-0.40)^2=0.545\]

@@

@@

@@

What's the probability that Gr\^emio looses: $p_{3h}=0.167$ or $p_{1a}=0.25$?

@@

@@

>>>>

\usepackage{amssymb}

@@

%More precisely, a scoring rule is a function of two variables (or vectors, matrices etc) that assigns a real number (score) to each possible outcome $X$ and its respective prevision value, $P(X)$.
%It is maximum for each $x$ when $x=P(X)$, and is nonincreasing both as $x$ departs from $P(X)$ (for each $P(X)$) and as $P(X)$ departs from $x$ (for each $x$).
%Thus, the best score is achieved when knowledge of $X$ is asserted as the value of $X$ itself, whatever it might be. Most of the scoring rules also exhibit an arbitrary normalizing feature such as $S[x,P(X)]=0$ for each $x$.

%Although many functions satisfy the criteria for a scoring rule specified above, not every scoring rule can be regarded as providing a fair assessment of the assertion $P(X)$, that is, not all scoring rules encourage honesty and accuracy in assessing numerical values for their previsions.
%Those that do encourage a fair assessment are named \emph{proper scoring rules} \cite{lad}.

%Say a forecaster's uncertainty about $X$ is $P_0$. A proper scoring rule is a scoring rule $S$ such that the prevision value $P(X)$ that minimizes $E_{P_0}[S(X,P(X))]$ is $P_0$. 
%In other words, if one receives a reward $S(x,P(x))$ when $x$ happens and if $S$ is a proper scoring rule, then the highest expected reward is obtained by reporting one's  true uncertainty about $X$.
%The use of a proper scoring rule thus encourages the forecaster to be honest to maximize the expected reward.
%Moreover, this method leads automatically to an overall comparison between the outcomes of different personal evaluations concerned with the same totality of events.
%Therefore, the accumulated score may be considered a concrete measure of success of each forecaster or model.

More precisely, a scoring rule is a function of two variables (or vectors, matrices etc) that assigns a real number (score) to each possible outcome $X$ and its respective prevision value, $P_X$. That is, $P_X$ is the probability assigned to each possible
outcome $X$.
It is minimum for each $x$ when $P_X$ is such that $P_X(X=x)=1$, and is nondecreasing both as $x$ departs from $P_X$ (for each $P_X$) and as $P_X$ departs from $x$ (for each $x$).
Thus, the best score (the smaller value of $S$) is achieved when knowledge of $X$ is asserted as the value of $X$ itself, whatever it might be.\footnote{A scoring rule can also be defined to be such that a \emph{large}
value of $S$ indicates good predictions.} 
%Most of the scoring rules also exhibit an arbitrary normalizing feature such as $S(x,P_X)=0$ for each $x$.

Although many functions satisfy the criteria for a scoring rule specified above, not every scoring rule can be regarded as providing a fair assessment of the assertion $P_X$, that is, not all scoring rules encourage honesty and accuracy in assessing numerical values for their previsions.
Those that do encourage a fair assessment are named \emph{proper scoring rules} \cite{lad}.

Say a forecaster's uncertainty about $X$ is $P_0$. A proper scoring rule is a scoring rule $S$ such that the prevision value $P_X$ that minimizes $E_{P_0}[S(X,P_X)]$ is $P_0$. 
In other words, if one receives a penalty $S(x,P_X)$ when $x$ happens and if $S$ is a proper scoring rule, then the lowest expected penalty is obtained by reporting one's  true uncertainty about $X$.
The use of a proper scoring rule thus encourages the forecaster to be honest to minimize the expected penalty.
Moreover, this method leads automatically to an overall comparison between the outcomes of different personal evaluations concerned with the same totality of events.
Therefore, the accumulated score may be considered a concrete measure of success of each forecaster or model.

@@

Let $P_X$ be a probability assignment for $X$.
The Brier score is given by
$$S(X;P_X)= \sum_{i}\mathbb{I}(X=x_i)(1- P_X(x_i))^2+\sum_{i}\mathbb{I}(X\neq x_i)P_X^2(x_i),$$
where $\mathbb{I}$ is the indicator function.

@@

We interpret the Brier score in the case only three events can happen, the case of a football match.
The green surface in Figure \ref{fig:simplex} represents the 2-simplex, i.e. the set of points such that $p_1+p_2+p_3=1$ for non-negative values of $p_1$, $p_2$ and $p_3$.
The triangle representing the simplex has sides of length $\sqrt{2}$ and its height is $\sqrt{6}/2$.
Drawing a similar equilateral triangle with height 1 and sides $2\sqrt{3}/3$, it is possible to represent all the points of the simplex.
This new triangle is used to graphically display the predictions as internal points because the sum of the distances of every point inside it to each side, representing the probability of each event, is always one.

@@

\node[label={below:Home wins}] at (0,0) {};
\node[label={below:Draw}] at (1.1547,0) {};
\node[label={above:Home loses}] at (0.57735,1) {};

@@

\node[label={below:Home wins}] at (0,-0.05) {};
\node[label={below:Draw}] at (1.1547,-0.05) {};
\node[label={above:Home loses}] at (0.57735,1.05) {};

@@

The Brier score for the prediction $p=(0.25,0.35,0.40)$ when the home loses is therefore given by
$d^2=(0-0.25)^2+(0-0.35)^2+(1-0.40)^2=0.545$. If the prediction was
 $p=(0,0,1)$, the score would be 0 instead.

@@

The logarithmic  score is given by
$$S(X;P(X))=- \sum_{i}\mathbb{I}(X=x_i)\log(P(x_i)),$$

This is the negative log likelihood
of the event that occurred.

The logarithmic score for the prediction $p=(0.25,0.35,0.40)$ when the home loses is therefore given by
$-\log(0.4)\approx 0.91$. If the prediction was
 $p=(0,0,1)$, the score would be $0$ instead.

@@

The spherical score is given by
$$S(X;P(X))=- \frac{1}{\sqrt{\sum_i P^2(x_i)}}\sum_{i}\mathbb{I}(X=x_i)P(x_i),$$

This is the negative likelihood
of the event that occurred, normalized by the square-root of the assigned probabilities.

The logarithmic score for the prediction $p=(0.25,0.35,0.40)$ when the home loses is therefore given by
$-0.4/\sqrt{0.25^2+0.35^2+0.40^2} \approx -0.68$. If the prediction was
 $p=(0,0,1)$, the score would be $-1$ instead.

@@

What's the probability that Gr\^emio loses: $p_{3h}=0.167$ or $p_{1a}=0.25$?

@@

In order to have a deeper understatement about the probabilities given by each method,
Figure \ref{fig::probMandante} displays the estimated conditional probability that the home team wins given there is no draw.
All models assign higher probabilities to the home team, a fact that illustrates that
they capture \red{?the well known fact?} that it is easier for a team to win at home. 
Morever, the outputs from the methods introduced in this paper
are less dispersed than those from the other models used in the literature.

\begin{figure}[H]
  \centering
\includegraphics[page=11,scale=0.4]{futebolComparacaoModelosForPaper.pdf}
  \caption{Conditional probability that the home team wins given there is no draw. Horizontal line indicates a 50\%
  probability.}
  \label{fig::probMandante}
\end{figure}

@@

\bibitem{lad}
Lad, F. Operational Subjective Statistical Methods: a mathematical, philosophical, and historical introduction. New York: Wiley, 1996.
